---
title: "EDA"
output: html_notebook
---

## Libraries
```{r}
library(tidyverse)
library(gmodels)
```

## Loading Data and Summaries

We initially thought of considering the entire dataset (both train and test)

But that only slowed down the visualizing.

```{r}
train_df <- read.csv("../credit_card_default_train.csv")
train_df$NEXT_MONTH_DEFAULT <- as.factor(train_df$NEXT_MONTH_DEFAULT)
test_df <- read.csv("../credit_card_default_test.csv")
df <- train_df
print(nrow(df))
print(colnames(df))
print(str(df))
print(summary(df))

```

## Visualizing Numerical Data

```{r}
ggplot(data=gather(train_df,key,value,c("DUE_AMT_JULY","DUE_AMT_AUG",
                                        "DUE_AMT_SEP","DUE_AMT_OCT",
                                        "DUE_AMT_NOV","DUE_AMT_DEC")) %>% 
         mutate(key = factor(key,levels=c("DUE_AMT_JULY","DUE_AMT_AUG",
                                        "DUE_AMT_SEP","DUE_AMT_OCT",
                                        "DUE_AMT_NOV","DUE_AMT_DEC"))),
       aes(x = key,y = value))+
  geom_violin()
```

```{r}
ggplot(data=gather(train_df,key,value,c("PAID_AMT_JULY","PAID_AMT_AUG",
                                      "PAID_AMT_SEP","PAID_AMT_OCT",
                                      "PAID_AMT_NOV","PAID_AMT_DEC")) %>%
         mutate(key = factor(key,levels=c("PAID_AMT_JULY","PAID_AMT_AUG",
                                      "PAID_AMT_SEP","PAID_AMT_OCT",
                                      "PAID_AMT_NOV","PAID_AMT_DEC"))),
       aes(x = key,y = value))+
  geom_violin()
```

## Patterns of response variable


We were initially going to look into the behaviours of clients through out the 6 months with regards to their delays. But there are too many observations to see a pattern
```{r}
ggplot(data=gather(train_df,key,value,c("PAY_JULY","PAY_AUG",
                                        "PAY_SEP","PAY_OCT",
                                        "PAY_NOV","PAY_DEC")) %>%
         mutate(key = factor(key,levels =c("PAY_JULY","PAY_AUG",
                                        "PAY_SEP","PAY_OCT",
                                        "PAY_NOV","PAY_DEC")), value=as.factor(value)),
       aes(x=key,y=value,group=Client_ID))+
  geom_line(aes(color=NEXT_MONTH_DEFAULT),size=0.5)+
  scale_y_discrete(labels=as.character(seq(-2,9)))+
  theme_minimal()
```
Cross tabulation of Balance Limit with response variable
```{r}
blv <- train_df%>% 
  mutate(Balance_Limit_V1=trimws(Balance_Limit_V1)) %>% 
  subset(select=Balance_Limit_V1)
train_df$Balance_Limit_V1 <- factor(as.factor(blv$Balance_Limit_V1),
                          levels=c("100K","200K","300K","400K","500K","1M","1.5M","2.5M"))
balance_response_ct <- CrossTable(train_df$Balance_Limit_V1 ,
                               train_df$NEXT_MONTH_DEFAULT ,
                               prop.r= FALSE,
                               prop.c=TRUE,
                               prop.chisq=FALSE, 
                               chisq = FALSE,
                               prop.t=TRUE)
```
Plotting
```{r}
ggplot(data=as.data.frame(balance_response_ct$prop.tbl),
       aes(fill=y,y=Freq,x=x))+
  geom_bar(stat = "identity",position = "fill")+
  scale_y_continuous(labels = scales::percent_format())+
  coord_flip()+theme_minimal()+
  xlab("Balance Limit")+ylab("Percentages")
```


```{r}
train_df <- train_df %>% mutate(
  Balance_Credit_Limit_Numeric = case_when(
    Balance_Limit_V1 == "100K" ~ 100000,
    Balance_Limit_V1 == "200K" ~ 200000,
    Balance_Limit_V1 == "300K" ~ 300000,
    Balance_Limit_V1 == "400K" ~ 400000,
    Balance_Limit_V1 == "500K" ~ 500000,
    Balance_Limit_V1 == "1M" ~ 1000000,
    Balance_Limit_V1 == "1.5M" ~ 1500000,
    Balance_Limit_V1 == "2M" ~ 2000000,
    Balance_Limit_V1 == "2.5M" ~ 2500000
  )
) %>%
  mutate(
    Due_Credit_Lim_JULY=(DUE_AMT_JULY/Balance_Credit_Limit_Numeric) * 100,
    Due_Credit_Lim_AUG=(DUE_AMT_AUG/Balance_Credit_Limit_Numeric) * 100,
    Due_Credit_Lim_SEP=(DUE_AMT_SEP/Balance_Credit_Limit_Numeric) * 100,
    Due_Credit_Lim_OCT=(DUE_AMT_OCT/Balance_Credit_Limit_Numeric) * 100,
    Due_Credit_Lim_NOV=(DUE_AMT_NOV/Balance_Credit_Limit_Numeric) * 100,
    Due_Credit_Lim_DEC=(DUE_AMT_DEC/Balance_Credit_Limit_Numeric) * 100
  )
```

```{r}
ggplot(data=gather(train_df,key,value,c("Due_Credit_Lim_JULY","Due_Credit_Lim_AUG",
                                        "Due_Credit_Lim_SEP","Due_Credit_Lim_OCT",
                                        "Due_Credit_Lim_NOV","Due_Credit_Lim_DEC")) %>%
         mutate(key = factor(key,levels=c("Due_Credit_Lim_JULY","Due_Credit_Lim_AUG",
                                        "Due_Credit_Lim_SEP","Due_Credit_Lim_OCT",
                                        "Due_Credit_Lim_NOV","Due_Credit_Lim_DEC"))),
       aes(x=key,y=value))+
  geom_boxplot(aes(fill=NEXT_MONTH_DEFAULT))+
  coord_cartesian(ylim=c(0,200))+
  scale_x_discrete(labels=c("JULY","AUG","SEP","OCT","NOV","DEC"))
```

## End of Day 1